---
title: "Serve Scouting Report"
subtitle: "Cal Poly vs. `r params$opponent`"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: false
    self_contained: true
params:
  opponent: "UC Santa Barbara"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)

# Resolve the data directory, matching the lookup in shiny_app/app.R: `data/`
# sits at the project root, but a deployed Shiny bundle keeps its own copy under
# shiny_app/. Check both so the report renders against either layout.
data_dir <- local({
  candidates <- c("../data", "../shiny_app/data")
  found <- candidates[file.exists(file.path(candidates, "serve_quality.rds"))]
  if (!length(found)) {
    stop("Could not find serve_quality.rds. Looked in: ",
         paste(normalizePath(candidates, mustWork = FALSE), collapse = ", "),
         ". See README > Reproducing the Data.")
  }
  found[1]
})

# serve_quality.rds is a superset of serves_featured.rds — it carries every
# feature column plus the OOF predictions and quality_index — so one file backs
# both the descriptive rate tables and the quality leaderboard.
serves    <- readRDS(file.path(data_dir, "serve_quality.rds"))
quality   <- serves
val_metrics <- if (file.exists(file.path(data_dir, "validation_metrics.rds"))) {
  readRDS(file.path(data_dir, "validation_metrics.rds"))
} else list(auc = "N/A")

opponent  <- params$opponent
cal_poly  <- "Cal Poly"

# Cal Poly green
cp_green <- "#154734"
```

```{r validate-opponent, include=FALSE}
all_teams <- unique(c(serves$serve_team, serves$opp_team))
if (!opponent %in% all_teams) {
  stop(paste0("Opponent '", opponent, "' not found. Available teams:\n",
              paste(sort(all_teams), collapse = ", ")))
}
```

```{r latest-rates}
# Most recent prior rates per player (from their last recorded contest).
# Ordered by date_parsed rather than contestid — NCAA IDs are assigned
# sequentially but monotonicity is not guaranteed by the API contract.
contest_dates <- readRDS(file.path(data_dir, "big_west_contests.rds")) %>%
  mutate(date_parsed = as.Date(date, "%m/%d/%Y")) %>%
  select(contestid = contest, date_parsed)

latest_server_rates <- serves %>%
  left_join(contest_dates, by = "contestid") %>%
  arrange(date_parsed) %>%
  group_by(player, serve_team) %>%
  slice_tail(n = 1) %>%
  ungroup() %>%
  select(player, serve_team, prior_ace_rate, prior_error_rate,
         prior_fbk_rate, match_ace_rate, match_error_rate)

latest_receiver_rates <- serves %>%
  filter(in_play == 1, !is.na(receiver)) %>%
  left_join(contest_dates, by = "contestid") %>%
  arrange(date_parsed) %>%
  group_by(receiver, opp_team) %>%
  slice_tail(n = 1) %>%
  ungroup() %>%
  select(receiver, opp_team, receiver_prior_fbk_rate)

latest_opp_rates <- serves %>%
  left_join(contest_dates, by = "contestid") %>%
  arrange(date_parsed) %>%
  group_by(opp_team) %>%
  slice_tail(n = 1) %>%
  ungroup() %>%
  select(opp_team, opp_prior_fbk_rate, opp_prior_ace_rate, opp_prior_error_rate)
```

---

## Their Weakest Receivers — Serve Targets

Players ranked by historical FBK concession rate (lower = weaker passer = better target).
Minimum 20 receptions. Focus on the top of this list.

```{r receiver-targets}
use_model_rates <- isTRUE(val_metrics$auc_m3 >= 0.60)

receiver_targets <- serves %>%
  filter(in_play == 1, opp_team == opponent, !is.na(receiver)) %>%
  group_by(receiver) %>%
  summarise(
    receptions   = n(),
    raw_fbk_rate = mean(fbk_against),
    .groups      = "drop"
  ) %>%
  filter(receptions >= 20) %>%
  left_join(
    latest_receiver_rates %>% select(receiver, receiver_prior_fbk_rate),
    by = "receiver"
  ) %>%
  mutate(
    fbk_rate = round(
      if (use_model_rates) coalesce(receiver_prior_fbk_rate, raw_fbk_rate) else raw_fbk_rate,
      3
    )
  ) %>%
  arrange(fbk_rate) %>%
  mutate(
    Rank           = row_number(),
    Recommendation = case_when(
      fbk_rate <= quantile(fbk_rate, 0.33) ~ "Primary Target",
      fbk_rate <= quantile(fbk_rate, 0.67) ~ "Secondary Target",
      TRUE                                  ~ "Avoid"
    )
  ) %>%
  select(Rank, Player = receiver, Receptions = receptions, `FBK Rate` = fbk_rate, Recommendation)

kable(receiver_targets, booktabs = TRUE, linesep = "") %>%
  kable_styling(font_size = 10, full_width = FALSE, position = "left") %>%
  row_spec(which(receiver_targets$Recommendation == "Primary Target"),
           bold = TRUE, color = "white", background = cp_green) %>%
  row_spec(which(receiver_targets$Recommendation == "Avoid"),
           color = "gray")
```

\vspace{6pt}
*FBK Rate = `r if (use_model_rates) "model-adjusted prior rate (receiver_prior_fbk_rate)" else "raw historical rate"`. Lower = weaker passer.*

---

## Their Serving Threats — Who to Watch

Cal Poly's defensive preparation. Opponent servers ranked by Serve Quality Index (0–100, global scale).
Minimum 50 serves.

```{r serving-threats}
threats <- quality %>%
  filter(serve_team == opponent, !is.na(quality_index)) %>%
  group_by(player, serve_team) %>%
  summarise(
    Serves   = n(),
    Quality  = round(mean(quality_index)),
    `P(Ace)` = round(mean(p_ace_oof), 3),
    `P(Err)` = round(mean(p_error_oof), 3),
    `P(FBK | In Play)` = round(mean(p_fbk_oof), 3),
    .groups  = "drop"
  ) %>%
  filter(Serves >= 50) %>%
  arrange(desc(Quality)) %>%
  mutate(Rank = row_number()) %>%
  select(Rank, Player = player, Serves, Quality, `P(Ace)`, `P(Err)`, `P(FBK | In Play)`)

kable(threats, booktabs = TRUE, linesep = "") %>%
  kable_styling(font_size = 10, full_width = FALSE, position = "left") %>%
  row_spec(1:min(3, nrow(threats)), bold = TRUE)
```

```{r dual-role-note, results='asis'}
dual_role <- intersect(receiver_targets$Player, threats$Player)
if (length(dual_role) > 0) {
  cat(paste0(
    "*Note: **", paste(dual_role, collapse = "**, **"),
    "** appear", if (length(dual_role) == 1) "s" else "",
    " in both tables. When serving, ",
    if (length(dual_role) == 1) "this player is" else "these players are",
    " unavailable as reception targets.*"
  ))
}
```

---

\vspace{12pt}
\footnotesize
*Receiver FBK rates above are observed 2025 season rates, not model predictions.
This report deliberately stops at the descriptive level. An earlier version included a
predicted server x receiver "matchup quality" matrix; it was removed because M3
(P(FBK Against)) has a chronological holdout AUC of `r val_metrics$auc`, which is too
close to chance to support player-level serve targeting. See the Limitations section of
the project README.*
